\documentclass[a4paper,10pt,reqno]{article}
\usepackage{amsmath,amssymb,amsfonts}
\renewcommand{\baselinestretch}{1.4}


\headheight=0in
\headsep=0in
\footskip=15pt
\textheight=680pt
\topskip=0pt
\topmargin=-10pt



\vfuzz5pt % Don't report over-full v-boxes if over-edge is small
\hfuzz5pt % Don't report over-full h-boxes if over-edge is small
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\centering{\Large\textbf{  Ordinary Differential Equations - 10413181}}
\centering{Homework No. 9}

\begin{enumerate}
\item Given the following inhomogeneous second order equation
\[ y'' - y' - 2y = 2 e^{-t} \]
\begin{enumerate}
\item Solve by comparing coefficients. (\emph{Make an ansatz for a particular solution $\psi = \ldots.$ If something goes wrong initially, try solving the homogeneous problem first to identify the origin of your problems. Then try $\psi = t \cdot \ldots$})
\item Use variation of parameters to solve the same problem.
\item Finally, given one solution $y_1$ of the homogeneous problem, use reduction of order with the ansatz $y = y_1(t) v(t)$ to derive a first order equation for $v'(t) = u(t).$ 
\item Show that finding a particular solution to the related problem \[ y'' - y' - 2y = 2 e^{-t} + 1 \]
is equivalent to adding a particular solution of 
$ y'' - y' - 2y = 1 $ to a particular solution of $y'' - y' - 2y = 2 e^{-t}.$ \emph{You may use any of the above methods to determine the particular solutions.}

\end{enumerate}
\item In class, we considered the case of damped harmonic oscillators
\[ m y'' + \gamma y' + k y = 0 \]
where $m$ is mass, $\gamma$ a damping coefficient, and $k$ a spring constant, and $m, \, k > 0, \, \gamma \geq 0$. Assume a motor is attached to the damped spring which supplies an oscillatory force $F_0 \cos(\omega t),$ so that 
\[ m y'' + \gamma y' + k y = F_0 \cos(\omega t). \]
\begin{enumerate}
\item We would like to find a particular solution to this general equation. Think first about which method is appropriate, and then use that method to obtain such a solution.
\item Look up the form of the homogeneous solutions from class (or derive them yourself). What is the long--time behavior ($t \rightarrow \infty$) of the general solution?
\item (Optional) Analyze the above problem for $\gamma = 0.$ Does any particular case of $\omega$ in the forcing play a special role? (\emph{We will discuss this in class.})
\end{enumerate}
\end{enumerate}

\end{document}


